Implementing simplified INLA into `aghq`

### 2022-10-13 src/dev_sinla/sinla.Rmd
Working on SINLA algorithm

# Marginal of joint Gaussian approximation

@@

The Gaussian approximation $p_\text{G}(x \, | \, \theta, y)$ is previously used in computing the Laplace approximation to $p(\theta \, | \, y)$
$$
\begin{equation}
\tilde p(\theta \, | \, y) \propto \frac{p(x, \theta, y)}{p_\text{G}(x \, | \, \theta, y)} \Big\rvert_{x = \hat{\mu}(\theta)} = \frac{p(y, \hat{\mu}(\theta), \theta)}{\det \left(\frac{1}{2 \pi} \hat{Q}(\theta) \right)^{1/2}}, (\#eq:pthetagiveny)
\end{equation}
$$
such that one easy way to calculate it is to use set `random = "x"` in the `TMB` template to integrate the latent field out, then recover the Gaussian approximation from there, as we have done in the function `h` above.
Let's optimise `h`:

@@

# Full Laplace approximation

@@

Another approach we might take is to calculate the full Laplace approximation
$$
\begin{align*}
  \tilde p(x_i \, | \, \theta, y) &\propto \frac{p(x, \theta, y)}{p_\text{G}(x_{-i} \, | \, x_i, \theta, y)} \Big\rvert_{x_{-i} = \hat{\mu}_{-i}(x_i, \theta)} \\
  &= \frac{p(y, x_i, \hat{\mu}_{-i}(x_i, \theta), \theta)}{\det(\frac{1}{2 \pi} \hat{Q}_{-i}(x_i, \theta))^{1/2}}
\end{align*}
$$
This is not practical in some settings as it involves recomputing the Gaussian approximation $p_\text{G}(x_{-i} \, | \, x_i, \theta, y)$ for each value of $(x_i, \theta)$ which can be too computationally expensive.
With of a bit of effort, we can implement this in `TMB` by passing in data where elements $i$ and $-i$ of the latent field are named as such (see Appendix for template).

@@

This remains to-do!

>>>>

# Marginal of joint Gaussian approximation {#gaussian-marginal}

@@

The Gaussian approximation $p_\text{G}(x \, | \, \theta, y)$ is previously used in computing the Laplace approximation to $p(\theta \, | \, y)$
$$
\begin{equation}
\tilde p(\theta \, | \, y) \propto \frac{p(x, \theta, y)}{p_\text{G}(x \, | \, \theta, y)} \Big\rvert_{x = \mu(\theta)} = \frac{p(y, \mu(\theta), \theta)}{\det \left(\frac{1}{2 \pi} Q(\theta) \right)^{1/2}}, (\#eq:pthetagiveny)
\end{equation}
$$
such that one easy way to calculate it is to use set `random = "x"` in the `TMB` template to integrate the latent field out, then recover the Gaussian approximation from there, as we have done in the function `h` above.
Let's optimise `h`:

@@

# Full Laplace approximation {#full-laplace}

@@

Another approach we might take is to calculate the full Laplace approximation
$$
\begin{align*}
  \tilde p(x_i \, | \, \theta, y) &\propto \frac{p(x, \theta, y)}{p_\text{G}(x_{-i} \, | \, x_i, \theta, y)} \Big\rvert_{x_{-i} = \mu_{-i}(x_i, \theta)} (\#eq:pxigiventhetay) \\
  &= \frac{p(y, x_i, \mu_{-i}(x_i, \theta), \theta)}{\det(\frac{1}{2 \pi} Q_{-i}(x_i, \theta))^{1/2}}
\end{align*}
$$
This is not practical in some settings as it involves recomputing the Gaussian approximation $p_\text{G}(x_{-i} \, | \, x_i, \theta, y)$ for each value of $(x_i, \theta)$ which can be too computationally expensive.
With of a bit of effort, we can implement this in `TMB` by passing in data where elements $i$ and $-i$ of the latent field are named as such (see Appendix for template).

@@

First, notational differences between Wood (2019) (former) and here (later) are the following:

* $\beta$ vs. $x$ for random effects
* $\hat \beta$ vs $\mu (\theta)$ for the random effects which maximise the template
* $\tilde \beta$ vs $(x_i, \mu(x_i, \theta))$ for the random effects which maximise the template subject to $x_i$ being fixed
* $H$ vs $Q(\theta)$ for the Hessian / precision matrix of the negative log-likelihood with respect to $x$ at $\mu(\theta)$

In Section \@ref(full-laplace) we directly calculate the Laplace approximation to $\int p(x, y, \theta) \text{d} \theta_{-i}$.
Following Wood (2019), a cheaper approach would be to "base the denominator of Equation \@ref(eq:pxigiventhetay) on the conditional density implied by ? in which case the Hessian is constant and
$$
\mu(x_i, \theta) = \mu(\theta) + \Sigma(\theta)_{-i, i} \Sigma(\theta)^{-1}_{i, i} (x_i - \mu (\theta)_i) (\#eq:wood3)
$$
The proposed modification has two parts:

1. Use the numerically exact $\mu(x_i, \theta)$ rather than an approximation (as in simplified Laplace)
2. An approximation to the required Hessian

## Complete algorithm

In this section we omit dependence on $\theta$ such that: $\mu = \mu(\theta)$, $Q = Q(\theta)$, $\mu(x_i) = \mu(x_i, \theta)$ and $Q(x_i) = Q(x_i, \theta)$.

Want: $\tilde p(x_i \, | \, \theta, y)$

Let $\mathcal{D}_{2j}$ denote a $2j \times 2j$ diagonal matrix with leading diagonal $-1, 1, -1, 1, \ldots$.
Let $u_0$ be a zero column matrix.
Let $\Delta_j^i$ denote $\Delta_j$ with an extra zero inserted at element $i$.

1. Compute the Cholesky factor of $Q_0 := Q_{-i, -i}$ by update of the Cholesky factor of $Q$
2. For each $x_i$ in a grid of evaluation points (how does one chose this grid?) repeat the following steps a-e)
  a) Use Newton's method with fixed Hessian $Q_{-i, -i}$ to find $\mu(x_i)$ starting from \@ref(eq:wood3)
  b) Compute a set of $J$ steps $\{\Delta_j\}$
  c) For $j = 1, \ldots, J - 1$ compute
  $$
  q = Q_0 \Delta_j + u_j \mathcal{D}_{2j}u_j^\top \Delta_j
  $$
  and
  $$
  g = \Delta_x \log p(\mu(x_i) + \Delta_j^i, y, \theta).
  $$
  Then compute the matrix
  $$
  u_{j + 1} = \{q(\Delta_j^\top j)^{-1/2}, g_{-i}(\Delta_j^\top g_{-i})^{-1/2}, u_j\}.
  $$
  d) Compute the determinant approximation...
  e) Compute $\tilde p(x_i \, | \, \theta, y) = \ldots$
3. Renormalise $\tilde p(x_i \, | \, \theta, y)$
